### 2020-01-24 thesis/start/english.tex
template adaptation

\iffalse \bibliography{../references.bib} \fi

%\begin{abstract}
your abstract in english
%\end{abstract}