### 2020-01-24 thesis/start/german.tex

\iffalse \bibliography{../references.bib} \fi

\begin{otherlanguage}{ngerman}

%\begin{abstract}
Deine Zusammenfassung in Deutsch.
%\end{abstract}

\end{otherlanguage}